The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
Further tune-ups
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@648 4525493e-7705-40b1-a816-d608a930855b

\documentclass[a4paper]{jpconf}
\usepackage{graphicx}
\begin{document}
\title{The CMS Data Aggregation System}

\author{Valentin Kuznetsov}
\address{Cornell University, Ithaca, New York, USA}
\ead{vkuznet@gmail.com}

\author{Dave Evans}
\address{Fermilab, Batavia, Illinois, USA}
\ead{evansde@fnal.gov}

\author{Simon Metson}
\address{Bristol University, Bristol, UK}
\ead{s.metson@bristol.ac.uk}

\begin{abstract}
Meta-data plays signiﬁcant role in large modern enterprises, 
research experiments and digital libraries where it comes from a different 
sources and distributed in a variety of forms and digital formats. 
It is organized and managed by constantly evolving software using 
both relational and non-relational data sources. Even though we can apply
information retrieval approach to non-relation data sources,
we can't do so for relational ones, where information can be accessed via
pre-established set of data-services.

Here we discuss a new data aggregation system which consumes, 
indexes and delivers information from different relational and 
non-relational data sources to answer cross data-service queries 
and explore meta-data associated with petabytes of experimental data. 
We combine simplpicity of keyword-based search with precision of RDMS
under the new system. The aggregated information being collected from various sources,
allowing end-users to place dynamic queries, get precise answers and 
trigger information retrieval on demand. Such close to real-time system 
based on studies  of use cases of the CMS particle physics experiment at 
the LHC where we used a large scale, distributed computing system 
to motivate this work.

\end{abstract}

\newpage

\section{Introduction}
The European Organization for Nuclear Research, known as CERN, plays a leading
role in fundamental studies of physics. But it is also known as a place where
many innovations in the area of computer science were developed, e.g. World Wide Web.
Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
Physics (HEP), promising to deliver a few PB of data each year. 
At this scale the ``information discovery'' within heterogeneous, distributed 
environment becomes a key ingredient of successful data analysis in HEP.
The data and associated meta-data are produced in variety of forms and digital formats.
They are stored and retrieved from relational and non-releational data-sources, such as 
RDMS systems, document oriented databases, blogs, twikies, file systems,
customized applications, etc. 
Working in such environment requires a lot of expertise and our users, physicists, 
are always looking for simple, intuitive and flexible
tool(s) to look-up their desired data. A well-known solutions, such as data-services
and search engines are tighted to a specific data sources and end-users are left 
with a manual task to bookkeep and relatate information from them.

@@

\section{Related Work\label{RelatedWork}}
Even though the idea of keyword queries over the relational database(s)
%querying relation databases via keyword based search
%algorithms 
is not knew it is still under discussion in computer science domain.
A few alternative solutions has been proposed to address this issue
in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
i.e. retrieval of only documents that contain all query keywords, has
been discussed. It requires that all keywords are matched in a row from a
single table or joined tables. Authors of \cite{QueryAnswer} made a step forward
by considering information around matched values and presents customized answers to
end-users over the generated new database. These and other related 
works\footnote{For more information about keyword based search over relational 
databases we redirect users to references in \cite{DBXplorer, QueryAnswer}.} 
are based on some kind of the graph over the underlying database and
moreover use a generated database out of the pristine one.
The proximity of the results in those techniques 
are mostly confined to the text-based values stored in a database, since 
usage of numerical values in the input search field is mostly useless without 
knowledge of the context of the value. For example, typing a value 100 
in the search input field leads to plenty of inappropriate matches, 
such as row ids, and in order to be
answered correctly requires additional keyword, clarifying its meaning, 
which should be matched with particular
entity (table.column) in the underlying database.
Therefore to answer presicely the logical questions, such as  
{\it find me total number of files whose size more then 20 but less then 
100} would be hard or almost impossible using keyword search techniques discussed 
in \cite{DBXplorer, QueryAnswer}, while it can be easily accomplished
using SQL syntax.

@@

At the same time the data aggregation is commonly addressed in the context of
federated databases, see for example \cite{FedDB}. It unifies data coming 
from different RDMS systems into the federated DB where SQL queries can be placed to 
search for desired data. Even though it would be interesting to apply technique 
discussed in \cite{DBS-QL} to a federated database it still requires 
an access to the underlying database schema. This obviously creates a problem
in distributed heterogeneous environment where different security policies
are in place.

To avoid those limitations we decided to build a keyword-search based system around
existing data-services whose nature and access policies are defined by developers.
Keeping in mind that meta-data can come in different forms and data formats
we used a document-oriented database back-end with full support of the query language.
The key ingredients were the setup of the mapping and analytics services which provide
necessary glue across existing API/notations and user queries. This approach 
allows us to perform the dynamic data aggregation across various data-services, achieve
precise answers using keyword-based queries
and organize information retrieval on demand.
%to store meta-data information 
%We wanted to expand this approach further and apply it 
%to handle cross data-service queries from broad variety 
%of relational and non-relational data-sources within 
%our distributed heterogeneous environment.

\section{CMS data model\label{DataModel}}
The CMS, Compact Muon Solenoid, \cite{CMS} 
is one of the two large general-purpose particle physics detectors built on 
the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
It is designed to explore the frontier of the High Energy Physics and provide physicists
ability to look at conditions presented at the early stage of our Universe.
More then 3000 physicists from 183 institutions reprepresenting almost 
38 countries are involved in the design, construction and maintenance of the experiment.

The CMS distributed computing and data model \cite{CMSDataModel} 
is designed to process
and efficiently manage a few PBs of data expected each year
in CMS operation at LHC. The computing resources provided by members of CMS
collaboration are geographically distributed, 
interconnected via high throughput networks and operated by means 
of Grid software. The model allows to cover broad variety of
hardware, mass and storage elements and configuration of the
clusters. To accomodate efficient data processing CMS uses
a multi-Tier distributed model, where specific tasks of data taking,
processing, archival and distribution are assigned to each tier based
on CMS data model. For example, the Tier-0 center at CERN is responsible
for archiving the data coming out from the detector, prompt first pass reconstruction
and data distribution to Tier-1 centers. The 7 Tier-1 centers
located in France, Germany, Ialy, Spain, Taiwan, United Kingdom and United States
keep a portion (copy) of the data delivered by Tier-0 for further processing.
They provide storage and CPU power for high priority analysis jobs.
The Tier-2 centers, located at more then 50 sites around the world,
are dedicated for user analysis tasks and production of simulated data.

A broad variety of data-services being designed and developed to
maintain detector and production operations, including detector
conditions databases, data location and bookkeeping services,
data transfer and job monitoring tasks. Even though majority of them
are located at CERN, it was never been a requirement in CMS computing
and data model. For instance, the production teams operated at Tier-1,2
centers set up local services for data bookkeeping and operational
tasks. Based on their resources the choice of database back-end was
up to the site managers. Therefore the CMS software were designed to support different
DB back-ends and provide tools for migrating data across them.

@@

\section{Data Aggregation System\label{DAS}}
The goal of the Data Aggregation System was to combine existing data-services
under one umrella, utilizing their access rules, APIs, security policies, etc., and
allowing users to perform queries against them. In contrast, approaches
discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
enforced a publication step of data into a dedicated database 
or central repositories \cite{iRODS}. Quite often it is required
that data should be converted into uniform data formats in order enable
aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
by existing data-services and infrastructure, whose design, implementation and
security policies are left up to their developers. Therefore DAS does not requires
a data preservation and transaction capabilities. Instead DAS can be treated
as proxy or caching layer where information retrieval (data) are triggered dynamically
by user queries. Such infrastructure allows data sharing across different queries and
data aggregation using appropriate mapping. Below we discuss details of
DAS architecture.

\subsection{DAS architecture}
The design of the DAS system was based on previous studies of CMS Data 
Bookkeeping System (DBS) \cite{DBS, DBS07}. We carefully analyzed user
requests and queries placed into the DBS, their patterns, frequencies and latency. 
As a result we created a data discovery service \cite{DD}. Originally it used MVC
architecture and later, being re-factored as pure presentation layer on top of 
the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
we took away relations and details on underlying schema from end-users. 
This was achieved by using Dijkstras shortest path algorithm 
to establish necessary table joins based on provided information in user queries and
knowledge of foreign key relationships from auto-loaded database schema.
The DBS-QL keys exposed to end-users were choosen based on a jargon used by
physicists. They're internally mapped into database entitites and attributes, 
e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
A quick adoption and wide usage of DBS-QL gave us confidence in choosen approach 
and provide a basis for building DAS system.

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support of keyword based search queries with ability to use condition operators;
\item pluggable interface to existing and up-coming data-service APIs;
\item support heterogeneous software environment and distributed nature of the data-services;
\item preserve security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementations and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=120mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query where placed into analytics DB and mapped into set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache users were able 
to see the results. This on demand system was supplemented by DAS robots 
(unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

\noindent
Due to this set of requirements, we rule out the choice of RDMS as DAS back-end(s)
for several reasons. We agreed to support dynamic queries, data
aggregation on demand and pluggable data-services. Therefore a choice 
of concrete schema for data storage was problematic. Moreover we didn't impose
any transactions and data persistency in DAS. This allows us to look 
for alternative IT solutions.
Through the analysis of available options, such as a file-based and memory caches, 
key-value databases, documented-oriented databases we made our choice in favor 
of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
MongoDB \cite{MongoDB}. Both systems provide schema free document
storage, replication and fail-over features, but our choice in favor of 
MongoDB was obvious, due to its support of dynamic queries, 
full indexies, including inner objects and embedded arrays,
and auto-sharding. Our prelimiary benchmarks shown that it can sustain
the desired load and size for our meta-data information. We used 
document-oriented database, MongoDB, as a back-end for the three 
main DAS components: Mapping, Analytics and DAS cache databases, 
which will be discussed next.

The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of the
following components:
core library with support of pluggable modules for data retrieval;
caching layer to store data-service output and aggregated results;
request service to handle user requests;
mapping DB to keep information about data-service APIs, their
notations and mapping to DAS;
analytics DB for query analysis and identification of pre-fetch 
strategies.

\subsection{DAS core}
The DAS core communicates with different data-services and retrieves
data on demand. This was done by invoking appropriate data-service plugin
upon the user request. Each plugin consists of DAS map,
data parser(s) suitable for data-service output format(s) into DAS common
data format (JSON) and was responsible for data-service API calls.
The data-service DAS map contains information about data-service APIs,
input and output parameters and their mapping to DAS keys. 
This information was stored into DAS Mapping DB which we discuss next.
Below you can see a typical DAS Mapping DB record for CMS
SiteDB data-service \cite{SiteDB}\footnote{We used regular expression 
patterns, evaluated during run time, to validate provided set of input 
paremters and DAS keys.}:
\begin{verbatim}
{"api": {"params": {"name": ""}, "name": "CMSNametoSE"}, 
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", "das_key": "site"}], 
 "_id": "4ae1f8dfe2194e4581000014", 
 "system": "sitedb", 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", "map": "site.name"}]}
\end{verbatim}
Although the majority of data-services looks alike
and their logic was abstracted, including parsers, there were a few
cases where special care was required. For instance,
we dealt with data-services which were only accessible via private network.
Therefore, a way to propagate user credentials was required, leading
to specific plugin implementation.

@@

\subsection{DAS Mapping DB}
DAS Mapping DB holded the mapping between data-service APIs and DAS keys exposed to
the users. It created a necessary glue between data-services and identifies
how their data can be related to each other. Basically Mapping DB was positioned 
as a {\it registration} service with ability to translate notations among data-services.
Upon identification of APIs which will participate in DAS activity 
we stored their names, input parameters, together with their type and accept 
values as regular expression patterns into the Mapping DB. 
Also we stored the mapping between data-service notations and DAS keys. 
For instance, the CMS DBS system \cite{DBS}
used {\it logical\_file\_name} notation to identify the name of the
logical file, while the same entity was called {\it lfn} in CMS PhEDEx system \cite{PhEDEx}.
Since both of them were referred to the logical file name, 
we map both notations into DAS key {\it file}.\footnote{We
also deal with physical file names, but still used {\it file} to refer to them in
DAS notations, since simple regular expressions were used to identify the
meaning of the {\it file} keyword from the provided value.}
Such keys, e.g. {\it file}, were exposed to the end-users as pre-defined keywords to
be used in search queries. For instance, end-users were able to type {\it file} in
order to get all file names known to DAS via data-services as well as
provide a concrete patterns associated with a file, e.g. {\it file=abc*}.
Under the hood, the appropriate data-service APIs were looked up along with appropriate
parameters to answer those questions. For example,
in aforementioned example, using {\it file} in DAS query cause call of DBS 
{\it listFiles} and PhEDEx {\it fileReplicas} APIs, respectively.
This approach gave power to end-users to express the meaning of their queries, especially
when they deal with numbers and achieve accuracy of results from DAS. 
For instance, users were able to type $run>10\,\,and\,\, run < 100$
to specify a range of run numbers to be looked up in DAS cache. At the same time
a full free text based keyword search is under development within MongoDB.

@@

The individual tests of DAS components (DBS and PhEDEx) shown roughly 
the same performance. Although the time of PhEDEx component was bigger 
due to longer retrieval and parsing time.
The time of PhEDEx component shown in Table \ref{DAS_benchmark} divided 
among: retrieval + parsing time $\sim$110 sec and merging step $\sim$50 sec.

>>>>

% Template for Elsevier CRC journal article
% version 1.0 dated 13 October 2009

% This file (c) 2009 Elsevier Ltd.  Modifications may be freely made,
% provided the edited file is saved under a different name

% This file contains modifications for Procedia Computer Science

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This template uses the elsarticle.cls document class and the extension package ecrc.sty
%% For full documentation on usage of elsarticle.cls, consult the documentation "elsdoc.pdf"
%% Further resources available at http://www.elsevier.com/latex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The '1p' and 'times' class options of elsarticle are used for Elsevier CRC
\documentclass[1p,times]{elsarticle}

%% The `ecrc' package must be called to make the CRC functionality available
\usepackage{ecrc}

%% The ecrc package defines commands needed for running heads and logos.
%% For running heads, you can set the journal name, the volume, the starting page and the authors

%% set the volume if you know. Otherwise `00'
\volume{00}

%% set the starting page if not 1
\firstpage{1}

%% Give the name of the journal
\journalname{Procedia Computer Science}

%% Give the author list to appear in the running head
%% Example \runauth{C.V. Radhakrishnan et al.}
\runauth{}

%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo

%% Give the abbreviation of the Journal.
\jid{procs}

%% Give a short journal name for the dummy logo (if needed)
\jnltitlelogo{Procedia Computer Science}

%% Hereafter the template follows `elsarticle'.
%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.

%% Elsevier CRC generally uses a numbered reference style
%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
%% If using BibTeX, use the style file elsarticle-num.bst

%% End of ecrc-specific commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

% if you have landscape tables
\usepackage[figuresright]{rotating}

% put your own definitions here:
%   \newcommand{\cZ}{\cal{Z}}
%   \newtheorem{def}{Definition}[section]
%   ...

% add words to TeX's hyphenation exception list
%\hyphenation{author another created financial paper re-commend-ed Post-Script}

% declarations for front matter

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\dochead{}
%% Use \dochead if there is an article header, e.g. \dochead{Short communication}

\title{The CMS Data Aggregation System}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

%\author[vkuznet]{Valentin Kuznetsov\corref{cor1}}
\author[vkuznet]{Valentin Kuznetsov}
\address[vkuznet]{Cornell University, Ithaca, New York, USA}
\ead{vkuznet@gmail.com}

\author[evans]{Dave Evans}
\address[evans]{Fermilab, Batavia, Illinois, USA}
\ead{evansde@fnal.gov}

\author[metson]{Simon Metson}
\address[metson]{Bristol University, Bristol, UK}
\ead{s.metson@bristol.ac.uk}

%\cortext[cor1]{Corresponding autho}

\begin{abstract}
Meta-data plays significant role in large modern enterprises, 
research experiments and digital libraries where it comes from a different 
sources and distributed in a variety of forms and digital formats. 
It is organized and managed by constantly evolving software using 
both relational and non-relational data sources. Even though we can apply
information retrieval approach to non-relation data sources,
we can't do so for relational ones, where information can be accessed via
pre-established set of data-services.

Here we discuss a new data aggregation system which consumes, 
indexes and delivers information from different relational and 
non-relational data sources to answer cross data-service queries 
and explore meta-data associated with petabytes of experimental data. 
We combine simplicity of keyword-based search with precision of RDMS
under the new system. The aggregated information being collected from various sources,
allowing end-users to place dynamic queries, get precise answers and 
trigger information retrieval on demand. Such close to real-time system 
based on studies  of use cases of the CMS particle physics experiment at 
the LHC where we used a large scale, distributed computing system 
to motivate this work.
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Meta-data\sep data aggregation\sep HEP.

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
% \linenumbers

%% main text
\section{Introduction}
The European Organization for Nuclear Research, known as CERN, plays a leading
role in fundamental studies of physics. But it is also known as a place where
many innovations in the area of computer science were developed, e.g. World Wide Web.
Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
Physics (HEP), promising to deliver a few PB of data each year. 
At this scale the ``information discovery'' within heterogeneous, distributed 
environment becomes a key ingredient of successful data analysis.
The data and associated meta-data are produced in variety of forms and digital formats.
They are stored and retrieved from relational and non-releational data-sources, such as 
RDMS systems, document oriented databases, blogs, twikies, file systems,
customized applications, etc. 
Working in such environment requires a lot of expertise and our users, physicists, 
are always looking for simple, intuitive and flexible
tool(s) to look-up their desired data. A well-known solutions, such as data-services
and search engines are tighted to a specific data sources and end-users are left 
with a manual task to bookkeep and relatate information from them.

@@

\section{Related Work\label{RelatedWork}}
Even though the idea of keyword queries over the relational database
%querying relation databases via keyword based search
%algorithms 
is not knew it is still under discussion in computer science domain.
A few alternative solutions has been proposed to address this issue
in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
i.e. retrieval of documents that contain all query keywords, has
been discussed. It requires that all keywords are matched in a row from a
single table or joined tables. Authors of \cite{QueryAnswer} made a step forward
by considering information around matched values and presents customized answers to
end-users over the generated new database. These and other related 
works, see \cite{DBXplorer, QueryAnswer} for more references,
are based on some kind of the graph over the underlying database and
use a generated database out of pristine one to perform the work.
The proximity of the results in those techniques 
are mostly confined to the text-based values stored in a database, since 
usage of numerical values in the input search field is mostly useless without 
knowledge of the context of the value. For example, typing a value 100 
in the search input field leads to plenty of inappropriate matches, 
such as row ids, and in order to be
answered correctly requires additional keyword, clarifying its meaning, 
which should be matched with particular
entity (table.column) in the underlying database.
Therefore to answer presicely the logical questions, such as  
{\it find me total number of files whose size more then 20 but less then 
100} would be hard or almost impossible in \cite{DBXplorer, QueryAnswer}, 
while it can be easily accomplished using SQL syntax.

@@

At the same time the data aggregation is commonly addressed in the context of
federated databases, see for example \cite{FedDB}. It unifies data coming 
from different RDMS systems into the federated DB where SQL queries can be user to 
search desired data. Even though it would be interesting to apply technique 
discussed in \cite{DBS-QL} to a federated DB it still requires 
an access to the underlying database schema. This obviously creates a problem
in distributed heterogeneous environment where different security policies
are in place.

To avoid those limitations we decided to build a keyword-search based system around
existing data-services whose nature and access policies are defined by developers.
Keeping in mind that meta-data can come in different forms and data formats,
we used a document-oriented database back-end with full support of the query language.
The key ingredients were the setup of the mapping and analytics services which provide
necessary glue across existing API/notations and user queries. This approach 
allows us to perform dynamic data aggregation across various data-services, achieve
precise answers using keyword-based queries
and organize information retrieval on demand.
%to store meta-data information 
%We wanted to expand this approach further and apply it 
%to handle cross data-service queries from broad variety 
%of relational and non-relational data-sources within 
%our distributed heterogeneous environment.

\section{CMS data model\label{DataModel}}
The CMS, Compact Muon Solenoid, \cite{CMS} 
is one of the two large general-purpose particle physics detectors built on 
the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
It is designed to explore the frontier of the HEP and provide physicists
the ability to look at conditions presented at the early stage of our Universe.
More then 3000 physicists from 183 institutions reprepresenting almost 
38 countries are involved in the design, construction and maintenance of the experiment.

The CMS distributed computing and data model \cite{CMSDataModel} 
is designed to process
and efficiently manage a few PBs of data expected each year
in CMS operation at LHC. The computing resources provided by members of CMS
collaboration are geographically distributed, 
interconnected via high throughput networks and operated by means 
of Grid software. The model allows to cover broad variety of
hardware, mass and storage elements and configuration of the
clusters. To accomodate efficient data processing CMS uses
a multi-Tier distributed model, where specific tasks of data taking,
processing, archival and distribution are assigned to each tier based
on CMS data model. For example, the Tier-0 center at CERN is responsible
for archiving the data coming out from the detector, prompt first pass reconstruction
and data distribution to 7 Tier-1 centers, located around the world.
%The 7 Tier-1 centers
%located in France, Germany, Ialy, Spain, Taiwan, United Kingdom and United States
%keep a portion (copy) of the data delivered by Tier-0 for further processing.
They provide storage and CPU power for high priority analysis jobs.
The 50+ Tier-2 centers
%, located at more then 50 sites around the world,
are dedicated for user analysis tasks and production of simulated data.

A broad variety of data-services being designed and developed to
maintain detector and production operations, including detector
conditions databases, data location and bookkeeping services,
data transfer and job monitoring tasks. Even though majority of them
are located at CERN, it was never been a requirement in CMS computing
and data model. 
%For instance, the production teams operated at Tier-1,2
%centers set up local services for data bookkeeping and operational
%tasks. Based on their resources the choice of database back-end was
%up to the site managers. Therefore the CMS software were designed to support different
%DB back-ends and provide tools for migrating data across them.

@@

\section{Data Aggregation System\label{DAS}}
The goal of the Data Aggregation System was to combine existing data-services
under one umrella, utilizing their access rules, APIs, security policies, etc., and
allowing users to perform queries against them. In contrast, the approaches
discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
enforced a publication step of data into a dedicated database 
or central repositories \cite{iRODS}. Quite often it is required
that data should be converted into uniform data formats in order enable
aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
by existing data-services and infrastructure, whose design, implementation and
security policies are left up to their developers. Therefore DAS does not require
a data preservation and transaction capabilities. Instead DAS can be treated
as proxy or caching layer where information retrieval are triggered dynamically
by user queries. Such infrastructure allows data sharing across different queries and
data aggregation using appropriate mapping. Below we discuss in details the
DAS architecture.

\subsection{DAS architecture}
The design of the DAS system was based on previous studies of CMS Data 
Bookkeeping System (DBS) \cite{DBS, DBS07}. Teh DBS Data Discovery service
\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provides
ability to place queries, similar to SQL, without specific knowledge of 
underlying DB schema.
%We carefully analyzed user
%requests and queries placed into the DBS, their patterns, frequencies and latency. 
%As a result we created a data discovery service \cite{DD}. Originally it used MVC
%architecture and later, being re-factored as pure presentation layer on top of 
%the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
%DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
%we took away relations and details on underlying schema from end-users. 
%This was achieved by using Dijkstras shortest path algorithm 
%to establish necessary table joins based on provided information in user queries and
%knowledge of foreign key relationships from auto-loaded database schema.
%The DBS-QL keys exposed to end-users were choosen based on a jargon used by
%physicists. They're internally mapped into database entitites and attributes, 
%e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
A quick adoption and wide usage of DBS-QL in CMS gave us confidence in choosen approach 
and provide a basis for building DAS system.

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support of keyword based search queries with ability to use condition operators;
\item pluggable interface to existing and up-coming data-service APIs;
\item support heterogeneous software environment and distributed nature of the data-services;
\item preserve security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementations and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query where placed into analytics DB and mapped into set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache users were able 
to see the results. This on demand system was supplemented by DAS robots 
(unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

\noindent
Based on this set we rule out the choice of RDMS as DAS back-end(s)
for several reasons. We agreed to support dynamic queries, data
aggregation on demand and pluggable data-services. Therefore a choice 
of concrete schema for data storage was problematic. Moreover we didn't impose
any transactions and data persistency in DAS. This allows us to look 
for alternative IT solutions.
Through the analysis of available options, such as a file-based and memory caches, 
key-value databases, documented-oriented databases we made our choice in favor 
of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
MongoDB \cite{MongoDB}. Both systems provide schema free document
storage, replication and fail-over features, but our choice in favor of 
MongoDB was obvious, due to its support of dynamic queries, 
full indexies, including inner objects and embedded arrays,
and auto-sharding. Our prelimiary benchmarks shown that it can sustain
the desired load and size of our meta-data information. We used 
document-oriented database, MongoDB, as a back-end for the three 
main DAS components: Mapping, Analytics and DAS cache databases, 
which will be discussed next.

The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of
core library with support of pluggable modules for data retrieval;
caching layer to store data-service output and aggregated results;
request service to handle user queries;
Mapping DB to keep information about data-service APIs, their
notations and mapping to DAS;
Analytics DB for query analysis and identification of pre-fetch 
strategies.

\subsection{DAS core}
The DAS core communicates with different data-services and retrieves
data on demand. This was done by invoking appropriate data-service plugin
upon the user request. Each plugin consists of DAS map, the
data parser(s) to convert data-service output into DAS records
and it was responsible for data-service API calls.
The data-service DAS map contains information about data-service APIs,
input and output parameters and their mapping to DAS keys. 
This information was stored into DAS Mapping DB which we discuss next.
Below you can see a typical DAS Mapping DB record for CMS
SiteDB data-service \cite{SiteDB}:
%\footnote{We used regular expression 
%patterns, evaluated during run time, to validate provided set of input 
%paremters and DAS keys.}:
\begin{verbatim}
{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", "das_key": "site"}], 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", "map": "site.name"}]}
\end{verbatim}
Although the majority of data-services looks alike
and their logic was abstracted, including parsers, there were a few
cases where special care was required. For instance,
to deal with data-services which were only accessible via private network.
Therefore, a way to propagate user credentials was required, leading
to specific plugin implementation.

@@

\subsection{DAS Mapping DB}
DAS Mapping DB holded the mapping between data-service APIs and DAS keys exposed to
the users. It created a necessary glue between data-services and identifies
how their data can be related to each other. Basically Mapping DB was positioned 
as a {\it registration} service with ability to translate notations among data-services.
Upon identification of APIs which will participate in DAS activity 
we stored their names, input parameters, together with their type and accept 
values as regular expression patterns into the Mapping DB. 
Also we stored the mapping between data-service notations and DAS keys. 
For instance, the CMS DBS system \cite{DBS}
used \verb+logical_file_name+ notation to identify the name of the
logical file, while the same entity was called \verb+lfn+ 
in CMS PhEDEx system \cite{PhEDEx}.
Since both of them were referred to the logical file name, 
we map both notations into DAS key \verb+file+.
%\footnote{We
%also deal with physical file names, but still used {\it file} to refer to them in
%DAS notations, since simple regular expressions were used to identify the
%meaning of the {\it file} keyword from the provided value.}
Such keys, e.g. \verb+file+, were exposed to the end-users as pre-defined keywords to
be used in search queries. For instance, end-users were able to type \verb+file+ in
order to get all file names known to DAS via data-services as well as
provide a concrete patterns associated with a file, e.g. \verb+file = abc*+.
Under the hood, the appropriate data-service APIs were looked up along with appropriate
parameters to answer those questions. In aforementioned example, 
usage of \verb+file+ in DAS query cause a call of DBS 
\verb+listFiles+ and PhEDEx \verb+fileReplicas+ APIs, respectively.
This approach gave power to end-users to express the meaning of their queries, especially
when they deal with numbers, and achieve accuracy of results. 
For instance, users were able to type \verb+run > 10 and run < 100+
to specify a range of run numbers to be looked up in DAS cache. At the same time
a full keyword search is under development.
% within MongoDB.

@@

The individual tests of DAS components (DBS and PhEDEx) shown roughly 
the same performance. Although the time of PhEDEx component was bigger 
due to longer retrieval and parsing time.
The time of PhEDEx component shown in Table \ref{DAS_benchmark} divided 
among: retrieval + parsing step, $\sim$110 sec, and merging step, $\sim$50 sec.
As shown in Table \ref{DAS_benchmark},
the time spent to access individual components (last column) was quite
reasonable, roughly 1 second for both systems. While final time to
get DAS records on disk was about 45 seconds. This was mostly due to I/O and
conversion operations between binary data format used by MongoDB, BSON, to DAS
data format, JSON.
